# Web Scraping Our Data

This section will describe the three different approaches I used to scrape datasets from Yelp, Goodreads, and MEC. 

thing reviewed, reviewer, date, text, and star rating.

## Foreword: Ethics & Legality

Web scraping is ubiquitous. As a rough working definition, let's say that web scraping--which can also be called crawling, trawling, harvesting, or any number of other terms--means automatically visiting websites to collect and store the information there so it can be used for a later purpose. So at one extreme, browsing Facebook at work doesn't count since it's not automatic; on the other extreme, automatically sending billions of requests to a server in an attempt to overload (i.e. a DDoS attack) doesn't count either, since nothing is being done with the information the server sends back.   To take the most obvious example, Google's entire business is based on @google-2020

It's also widespread--it's how Google gets its information,  something like 50% of page visits are for scraping? source this. There are many commercial web-scraping companies, and programming languages like R have web-scraping libraries. 

Lots of worrying about this

@scasa-2018 gives a good overview of the legal situation from a Canadian perspective, although it came out before the Mongohouse case.

[mongodb opinion piece](https://www.canadianlawyermag.com/news/opinion/federal-court-makes-clear-website-scraping-is-illegal/276128)

[Apparently it's legal!](https://www.forbes.com/sites/emmawoollacott/2019/09/10/linkedin-data-scraping-ruled-legal/#56172d721b54)



To mitigate any ethical concerns, in this project we've made the following choices:

* We're scraping a reasonably small number of pages/reviews;
* We're being considerate of their servers by spacing out our requests; and,
* We're collecting and using the data for educational non-commercial purposes.

## The General Idea

Web scraping these sites follows a two-step process:

1. Get a list of urls for pages you want to scrape (generating an *index*).
  * Usually we'll get these urls by first scraping another page.
1. Use a loop to scrape the information from each page (loading the *content*).

Since different sites have different structures, we'll need custom code for the index and content pages. Also, by random chance these three sites all use different web-design principles, so we'll also need to use different techniques. 


## Goodreads: CSS Selectors

[Goodreads](www.goodreads.com) describes itself as "the world’s largest site for readers and book recommendations" (@goodreads-2020). Registered users can leave reviews and ratings for books, any anyone can use the site to browse user-submitted reviews and a variety of information about books. 

Goodreads' pages are standard html, so we can use css selectors to isolate the exact parts of the page we're interested in. I used R's **rvest** package, and the package documentation has details about the methods and about css selectors in general [@wickham-package-2020]. To find the css selectors I used [SelectorGadget](https://selectorgadget.com/), a point-and-click Chrome extension.

### Scraping the Index

Goodreads assigns books to genres like "sci-fi" and "romance," and curates lists of each genre's 100 most-read books in the past week. By scraping these pages, we can get links to content pages for hundreds of books across different genres.

Here is a code block to get the links to the 100 most-read books in the "classics" genre. The code could be functionized or run several times for other genres.

```{r, eval=FALSE}
library(tidyverse)
library(rvest)

# choose a genre
genre <- "classics"

url <- paste0("https://www.goodreads.com/genres/most_read/",genre)

# read the page
page <- read_html(url)

# extract the links to each book's page using css selectors
book_links <- page %>%
  html_nodes(".coverWrapper") %>%
  html_nodes("a") %>%
  html_attr("href") %>%
  paste0("https://www.goodreads.com", .)

book_links %>%
  as_tibble() %>%
  write_csv(paste0("book_links_",genre,".csv"))
```


### Scraping the Content

The next step is to load each content link and extract the information about the book, its author, and all the reviews. Books can have several pages of reviews, so we need to figure out how mayn pages there are and how to crawl through them. Since not all reviews have both text and star ratings, we also need to be careful to make sure we handle missing data appropriately. 

```{r, eval=FALSE}

#https://www.goodreads.com/genres/most_read/non-fiction
links <- book_links

# set up empty results tibble
results <- tibble()

# remove any links we've already seen, if we crashed and are resuming
links <- links[!links %in% results$url]

for (i in 1:length(links)) {
  
  # pause briefly
  pause()
  
  # get the url we're interested in
  url <- links[[i]]
  
  
  # write an update, since I'm impatient and want to know what's happening
  message(paste0(i,"/",length(links),": ", url))
  
  # choose a random useragent each time we load the page -- anti-anti-scraping measure
  httr::user_agent(random_useragent()) %>%
    httr::set_config()
  # read the url
  
  page <- read_html(url)
  
  # read the review page's html
  reviews_html <- page %>%
    html_nodes(".review")
  
  # extract the informaiton we're interested in
  book_title <- page %>%
    html_nodes("#bookTitle") %>%
    html_text() %>%
    stringr::str_trim()
  
  author_name <- page %>%
    html_nodes(".authorName span") %>%
    html_text() %>% head(1)
  
  review_names <- purrr::map_chr(reviews_html, function(x) { html_nodes(x, ".user") %>% html_text() })
  review_dates <- purrr::map_chr(reviews_html, function(x) {html_nodes(x, ".reviewDate") %>% html_text()})
  review_text <- purrr::map_chr(reviews_html, function(x) {html_nodes(x, ".readable span") %>% html_text() %>% paste0(., " ") %>% na_if(y=" ") %>% str_trim() %>% tail(1)})
  review_rating <- purrr::map_chr(reviews_html, function(x) {html_nodes(x, ".staticStars") %>% html_text() %>% paste0(., " ") %>% na_if(y=" ") %>% str_trim()}) 
  
  # how many pages of reviews?
  # there may be an easier way but this should work
  num_pages <- page %>%
    html_text() %>%
    str_extract_all("(?<=previous).*?(?=next)") %>%
    unlist() %>% tail(1) %>%
    stringr::str_trim() %>%
    stringr::str_split(" ") %>%
    unlist() %>%
    map_dbl(as.double) %>%
    max()
  
  # put it all together
  page_reviews <- tibble(
    book_title = book_title,
    author_name = author_name,
    comment = review_text,
    names = review_names,
    rating = review_rating,
    dates = lubridate::mdy(review_dates),
    url = url,
    num_pages = num_pages
  ) 
  
  results <- bind_rows(results, page_reviews)
}

filename <- paste0("goodreads_",genre,"_reviews.csv")
results %>%
  write_csv(path = filename)
```


## Yelp: Embedded JSON

[Yelp](www.yelp.com), according to its website, "connects people with great local businesses" [@yelp-about-2020]. Businesses can upload information like their location, hours, and services, and registered users can can leave reviews with text, star ratings, and pictures. 

Yelp's web design includes structured json data within its html. In other words, each Yelp review page has machine-readable data hidden inside it if you know where to look. We'll exploit this by using a regular expression to extract the json from the html, then parse the json and work with it directly.

### Scraping the Index

First we'll get the urls for each restaurant in Ottawa. We start at the base url for Ottawa restaurants and iterate through all of the pages: we could get the page numbers automatically, but here I just saw that there are 24 and hard-coded that number in. We extract the urls from the json in the page *without* parsing it using a regex. We could have parsed the json and done it using structured data, but since we're only looking for one value type this was faster and worked fine.

```{r, eval=FALSE}
# the base url for restaurants in Ottawa
baseurl <- "https://www.yelp.ca/search?cflt=restaurants&find_loc=Ottawa%2C%20Ontario%2C%20CA"

# an empty tibble for our links
links <- tibble()

# loop through all 24 pages of Ottawa restaurants. (The number 24 was hard-coded to keep things moving.)
for (pagenum in 1:24){
  Sys.sleep(1) 
  
  # get the url for the page we're loading
  url <- paste0(baseurl, if(pagenum>1){ paste0("&start=",(pagenum-1)*10) })
  
  # load the html for the page and print an update message
  text <- read_html(url) %>%
    html_text() 
  message("**PAGE ",pagenum,": ", url)
  
  # extract the urls using a straight regex based on the json value key. we're not parsing any json here.
  urls <- text %>%
    str_extract_all('(?<=businessUrl":")(.*?)(?=")') %>%
    unlist() %>%
    enframe() %>%
    select(-name) %>%
    filter (!str_detect(value, "ad_business_id")) %>%
    distinct() %>%
    transmute(url = paste0("http://www.yelp.ca", value))
  
  # add to our results
  links <- bind_rows(links, urls)
}

links %>%
  write_csv("yelp_ottawa_links.csv")
```


### Scraping the Content

Scraping the content has two steps. First, now that we have a list of content urls, we can load each in turn and extract the reviews and the links for any additional review pages for this business. In the second step we'll load these new links and get those reviews. 

This function loads a single review page, extracts the machine-readable json using a regex, parses the json, and extracts the information we're interested in. It then returns that information in a tibble.

```{r function_get_review, eval=FALSE}
get_review <- function(page, url) {
  
  # get the html
  text <- page %>%
    html_text()
  
  # extract the json with the review data
  json_text <- text %>%
    str_extract('(?<="reviewFeedQueryProps":)(.*)("query":""\\}\\})')
  
  # set our review_page results variable to NA, in case we don't get a results
  review_page <- NA
  
  # make sure we have valid json text before we try to parse it
  if (!is.na(json_text)){
    # parse the json
    json_parse <- json_text %>%
      jsonlite::fromJSON()
    
    # pull out the variables we're interested in
    review_text <- json_parse$reviews$comment$text
    review_rating <- json_parse$reviews$rating
    review_name <- json_parse$reviews$user$markupDisplayName
    review_date <- json_parse$reviews$localizedDate
    review_business <- json_parse$reviews$business$name
    review_url <- rep(url, length(review_text))
    
    # put them all into a tibble
    review_page <- tibble(business = review_business,
                          name = review_name,
                          date = review_date,
                          comment = review_text,
                          rating = review_rating,
                          url = review_url)
  }
  
  # return either NA or a results tibble
  return (review_page)
}


# simple function to pause for a random period of time
pause <- function(min_wait = 1, max_wait = 3){
  runif(n=1, min=min_wait, max = max_wait) %>% Sys.sleep()
}

```

We then proceed with step one, loading the initial list of links, extracting the reviews there, and collecting any more links to more reviews:



```{r, eval=FALSE}
# load our set of restaurant page links
base_links <- read_csv("yelp_ottawa_links.csv")

# set up an empty tibble for our reviews
reviews <- tibble()

# set up an empty tibble for the links we're going to visit later
more_links <- tribble(~links)

# now we're going to visit each page, extract the reviews from it, and find out how many *more* pages there are for this restaurant.
# we'll keep track of those other pages and visit them later in a random order.
for (i in 1:nrow(base_links)) {
  # pause briefly
  pause()
  
  # get the url we're interested in
  url <- links[[i]]
  
  # write an update, since I'm impatient and want to know what's happening
  message(paste0(i,"/",nrow(base_links),": ", url))
  
  # read the url
  page <- read_html(url)
  
  # extract the reviews from the page
  review_page <- get_review(page, url)
  
  # add these reviews to our list of reviews
  reviews <- bind_rows(reviews, review_page)
  
  # now find out how many other pages there are for this restaurant
  # we'll regex to find the second half of "dd of dd", where d is a digit (and it could be either one or two digits--see the regex below)
  num_pages <- page %>%
    html_node((".text-align--center__373c0__2n2yQ .text-align--left__373c0__2XGa-")) %>%
    html_text() %>%
    str_extract("(?<=of )(\\d\\d?)") %>%
    as.integer()
  
  # make sure we don't get an NA
  if (is.na(num_pages)) num_pages <- 1
  
  # if there's more than one page, construct the links and add them to our list of links to read next
  if (num_pages > 1) {
    more_links <- more_links %>% 
      add_row(links = paste0(url, "?start=",(1:(num_pages-1))*20) )
  }
  
} # end for i in 1:nrow(base_links)

# save our results
reviews %>%
  write_csv("data/ottawa-reviews-1.csv")

more_links %>%
  write_csv("data/ottawa_more_links.csv")

```

In step two, we'll repeat the process for the new links we collected:


Now let's do the same thing for the extra links we got:
note it's stopping me every 136 or so and giving a 503 error, so i'm either rebooting my modem to get a new ip address or tethering to my phone for a bit

```{r, warning=FALSE, eval=FALSE}
links <- more_links

for (i in 1:length(links)) {
  # pause briefly for random interval
  pause()
  
  # get the url we're interested in
  url <- links[[i]]
  
  # write an update, since I'm impatient and want to know what's happening
  message(paste0(i,"/",length(links),": ", url))

  message("  Loading page.")
  # read the url
  page <- read_html(url)
  
  message("  Parsing review.")
  # extract the reviews from the page
  review_page <- get_review(page, url)
  
  if (!is.na(review_page)){
    message ("  Adding to inventory.")
    # add these reviews to our list of reviews
    reviews <- bind_rows(reviews, review_page)
  } else {
    message ("  No valid json found.")
  }
} # end for i in 1:nrow(base_links)


reviews %>%
  write_csv("ottawa-reviews-2.csv")

```

## MEC: Reverse-Engineering Client-Side API Calls

MEC's website uses a completely different design principle that makes it seem more difficult to extract information. If you inspect the html for one of MEC's product pages, you'll find that the review information simply isn't there! It's quite mysterious.

The secret is that MEC's site uses client-side API calls to download the data which is then displayed locally. To solve this puzzle, I needed to use  Chrome's developer console (opened with *Control-Shift-J*) to see the network activity (under the *Network* tab) happening each time I loaded a new product page. I discovered that my browser was making API calls to a specific server, and by comparing the calls for a few products I found that the main difference was the product ID. This let me reverse-engineer the syntax just enough to be able to call it myself and get reviews for any product based on its ID. I also found that there was one API call for the first page of reviews and a different one for loading more reviews, so I built functions for both of them.

As a result, the index in this case is a list of product IDs rather than urls, and the content is the result of API calls rather than web pages. However, the principles remain the same.

### Scraping the Index

This code block collects product IDs for [mittens and gloves](https://www.mec.ca/en/products/clothing/clothing-accessories/gloves-and-mittens/c/987). Each product category has a different catalogue page, so I modified the code to load a few different kinds of products. We load the first page, use a regex to figure out how many pages there are, then use css selectors to extract the IDs for products with reviews.

```{r, eval=FALSE}
# enter the base url by hand
base_url <- "https://www.mec.ca/en/products/clothing/clothing-accessories/gloves-and-mittens/c/987"

# enter the product type by hand
product_type <- "gloves-and-mittens"

# read the page
page <- read_html(base_url)

# get the number of items using a CSS selector and a regex
# we expect to find between one and three digits
num_items <- page %>%
  html_nodes(".qa-filter-group__count") %>%
  html_text() %>%
  str_extract("(\\d\\d?\\d?)") %>%
  as.integer()

# there are at most 36 items per page
num_pages <- (num_items / 36) %>% ceiling()

# first let's do the items on this page
# find each link to a product, filter out any that don't have reviews yet, extract the product ids
product_ids <- page %>%
  html_nodes(".rating__count__link") %>%
  html_attrs() %>%
  enframe() %>%
  unnest_wider(value) %>%
  filter(!str_detect(title, "No reviews yet")) %>%
  mutate(product_id = str_extract(href, "\\d\\d\\d\\d-\\d\\d\\d")) %>%
  select(-name, -class)

# now we load the extra pages, if there are any
if (num_pages > 1) {
  # we iterate from 1 to num_pages-1, because MEC calls the first extra page page 1
  for (i in 1:(num_pages-1)){
    # send an update to the console
    message(paste0(i,"/",(num_pages-1)))
    
    # wait a little bit
    patience(min_wait = 3, max_wait = 10)
    
    # get the new url for the next page
    url <- paste0(base_url,"?page=",i)
    
    # load the next page
    page <- read_html(base_url)
    
    # find each link to a product, filter out any that don't have reviews yet, extract the product ids
    new_product_ids <- page %>%
      html_nodes(".rating__count__link") %>%
      html_attrs() %>%
      enframe() %>%
      unnest_wider(value) %>%
      filter(!str_detect(title, "No reviews yet")) %>%
      mutate(product_id = str_extract(href, "\\d\\d\\d\\d-\\d\\d\\d")) %>%
      select(-name, -class)
    
    # add it to our list
    product_ids <- bind_rows(product_ids, new_product_ids)
    
  } # end for (i in 1:(num_pages-1))
} # end if (num_pages >1)


product_ids %>%
  write_csv(paste0("data/product_ids_",product_type,".csv"))

```


### Functions for API Calls  

Next, I defined functions to make the API calls and to process their results. The API calls are quite ugly--I could have spent more time figuring out exactly how they worked and slimmed them down, but this worked.

```{r, eval=FALSE}
get_first_api_url <- function(product_code){
  api_url <- paste0("https://api.bazaarvoice.com/data/batch.json?passkey=dm7fc6czngulvbz4o3ju0ld9f&apiversion=5.5&displaycode=9421-en_ca&resource.q0=products&filter.q0=id%3Aeq%3A",product_code,"&stats.q0=questions%2Creviews&filteredstats.q0=questions%2Creviews&filter_questions.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_answers.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_reviews.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_reviewcomments.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&resource.q1=questions&filter.q1=productid%3Aeq%3A",product_code,"&filter.q1=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q1=totalanswercount%3Adesc&stats.q1=questions&filteredstats.q1=questions&include.q1=authors%2Cproducts%2Canswers&filter_questions.q1=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_answers.q1=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q1=10&offset.q1=0&limit_answers.q1=10&resource.q2=reviews&filter.q2=isratingsonly%3Aeq%3Afalse&filter.q2=productid%3Aeq%3A",product_code,"&filter.q2=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q2=helpfulness%3Adesc%2Ctotalpositivefeedbackcount%3Adesc&stats.q2=reviews&filteredstats.q2=reviews&include.q2=authors%2Cproducts%2Ccomments&filter_reviews.q2=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_reviewcomments.q2=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_comments.q2=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q2=8&offset.q2=0&limit_comments.q2=3&resource.q3=reviews&filter.q3=productid%3Aeq%3A",product_code,"&filter.q3=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q3=1&resource.q4=reviews&filter.q4=productid%3Aeq%3A",product_code,"&filter.q4=isratingsonly%3Aeq%3Afalse&filter.q4=issyndicated%3Aeq%3Afalse&filter.q4=rating%3Agt%3A3&filter.q4=totalpositivefeedbackcount%3Agte%3A3&filter.q4=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q4=totalpositivefeedbackcount%3Adesc&include.q4=authors%2Creviews%2Cproducts&filter_reviews.q4=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q4=1&resource.q5=reviews&filter.q5=productid%3Aeq%3A",product_code,"&filter.q5=isratingsonly%3Aeq%3Afalse&filter.q5=issyndicated%3Aeq%3Afalse&filter.q5=rating%3Alte%3A3&filter.q5=totalpositivefeedbackcount%3Agte%3A3&filter.q5=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q5=totalpositivefeedbackcount%3Adesc&include.q5=authors%2Creviews%2Cproducts&filter_reviews.q5=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q5=1&callback=BV._internal.dataHandler0")

    return(api_url)
}

get_second_api_url <- function(product_code){
  api_url <- paste0("https://api.bazaarvoice.com/data/batch.json?passkey=dm7fc6czngulvbz4o3ju0ld9f&apiversion=5.5&displaycode=9421-en_ca&resource.q0=reviews&filter.q0=isratingsonly%3Aeq%3Afalse&filter.q0=productid%3Aeq%3A",product_code,"&filter.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q0=helpfulness%3Adesc%2Ctotalpositivefeedbackcount%3Adesc&stats.q0=reviews&filteredstats.q0=reviews&include.q0=authors%2Cproducts%2Ccomments&filter_reviews.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_reviewcomments.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_comments.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q0=30&offset.q0=8&limit_comments.q0=3&callback=bv_351_44883")
  
  #api_url <- paste0("https://api.bazaarvoice.com/data/batch.json?passkey=dm7fc6czngulvbz4o3ju0ld9f&apiversion=5.5&displaycode=9421-en_ca&resource.q0=reviews&filter.q0=isratingsonly%3Aeq%3Afalse&filter.q0=productid%3Aeq%3A",product_code,"&filter.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&sort.q0=helpfulness%3Adesc%2Ctotalpositivefeedbackcount%3Adesc&stats.q0=reviews&filteredstats.q0=reviews&include.q0=authors%2Cproducts%2Ccomments&filter_reviews.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_reviewcomments.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&filter_comments.q0=contentlocale%3Aeq%3Aen*%2Cfr_CA%2Cen_CA&limit.q0=500&offset.q0=0&limit_comments.q0=3&callback=bv_351_44883")
  
  return(api_url)
}

# function to go through the list and extract meaningful results
get_review <- function(x){
  product_id <- ifelse(!is.null(x$ProductId), x$ProductId, "")
  user_name <- ifelse(!is.null(x$UserNickname), x$UserNickname, "")
  rating <- ifelse(!is.null(x$Rating), x$Rating, 0)
  
  review_date <- ifelse(!is.null(x$SubmissionTime), x$SubmissionTime, "")
  review_text <- ifelse(!is.null(x$ReviewText), x$ReviewText, "")
  review_title <- ifelse(!is.null(x$Title), x$Title, "")
  
#  message ("sofa sogood")
  
  results <- tibble(
    product_id = product_id,
    user_name= user_name,
    rating_num = rating,
    review_date =review_date,
    review_text =review_text ,
    review_title = review_title
  )
  return(results)
}

```



### Scraping the Content

Now that we have the product ids, we can loop through them and call the API to get the reviews.

```{r, eval=FALSE}
# set up our results tibble
all_reviews <- tibble()

# #20 seems to have no reviews, json_results didn't have SubmissionTime, so added that to conditions
for (i in 1:nrow(product_ids)){
  # print an update message and wait nicely
  product_id <- product_ids$product_id[[i]]
  message(paste0("Product #",i,"/",nrow(product_ids),": ",product_id))
  Sys.sleep(2)
  
  api_url <-  get_first_api_url(product_id)
  
  # call the API
  text <- GET(api_url) %>%
    content("text")
  
  # parse the returned text into json
  json_parsed <- text %>%  
    str_extract("\\{(.*)\\}") %>%
    #str_extract("(?<=BV._internal.dataHandler0\\()(.*)")#(?=\\))") %>%
    jsonlite::parse_json()
  
  # get the product information
  product <-  json_parsed$BatchedResults$q0$Results[[1]]
  product_name <- product$Name
  product_brand <- product$Brand$Name
  
  
  reviews <- json_parsed$BatchedResults$q2$Results
  
  # use purrr::map to apply get_review() to each individual review
  reviews1 <- tibble(
    x = purrr::map(reviews, get_review)
  ) %>%
    unnest(cols = "x")
  
  message ("   First API call done and processed.")
  
  ####################################3
  # SECOND API CALL. Try to load additional reviews:
  api_url <-  get_second_api_url(product_id)
  # test <- read_html(api_url)
  # text <- test %>% html_text() 
  text <- GET(api_url) %>%
    content("text")
  
  json_parsed <- text %>%
    str_extract("(?<=\\()(.*)(?=\\))") %>%
    jsonlite::fromJSON()
  
  json_results <- json_parsed$BatchedResults$q0$Results 
  
  
  # set our second set of reviews to NULL in case we don't find any
  reviews2 <- NULL
  
  # if we do find some, set them to that!
  if (!is.null(json_results) & length(json_results)>0) {
    if (any(str_detect(names(json_results), "SubmissionTime"))){
      reviews2 <-   json_results %>%
        as_tibble() %>%
        select(review_date = SubmissionTime,
               user_name = UserNickname,
               review_title = Title,
               review_text = ReviewText,
               rating_num = Rating
        ) %>%
        mutate(product_id = product_code)
    }
  }
  
  message ("    Second API call done and processed.")
  
  # put the new reviews together:
  new_reviews <-  bind_rows(reviews1, reviews2) %>%
    mutate(product_name = product_name,
           product_brand= product_brand)
  
  all_reviews <- bind_rows(all_reviews, new_reviews)
} # end (for i)

all_reviews %>%
  distinct() %>%
  write_csv(paste0("reviews-",product_type,".csv"))
```

## Summary

